---
title: "Mini Project 4: Text Analysis"
format: html
---

## Introduction

For this project, I explored two very different texts:

-   **The Book of Nonsense** by Edward Lear, a collection of whimsical limericks filled with absurd characters and places (which was published in 1846!).

-   **The 2002 CIA World Factbook**, a straightforward encyclopedia of global country data, such as geography, economy, and demographics. I chose 2002 because I thought it would be especially interesting to see what the CIA had to say after 9/11.

By combining nonsense with writing from the CIA, I wanted to look at the language differences between the two different text.

*Getting the environment ready for the project:*

```{r, include = FALSE}
library(RCurl)
library(tidyverse)
library(tidytext)
library(textdata)
library(wordcloud)
library(wordcloud2)
library(viridis)
library(ggthemes)
library(gutenbergr)
library(RCurl)
library(wordcloud)


# Choose a lexicon
bing <- get_sentiments("bing")
nrc <- get_sentiments("nrc")




```

## Importing the data and doing some basic cleaning:

```{r}
#6344 = CIA 2002
#982 = The Book of Nonsense
Books <- gutenberg_download(c(6344, 982), meta_fields = "title", mirror = "http://mirror.csclub.uwaterloo.ca/gutenberg/")

write_csv(Books, "~/SDS264_F24/books.csv") #writes csv so we can use it


Books <- read_csv("~/SDS264_F24/books.csv", show_col_types = FALSE) #reads csv

Books |> count(title)
Books

nonsense <- Books |>
  filter(str_detect(title, "The Book of Nonsense")) #detects if the title matches.
CIA <- Books |>
  filter(str_detect(title, "The 2002 CIA World Factbook"))


Books |> count(title)
Books

nonsense <- Books |>
  filter(str_detect(title, "The Book of Nonsense")) #creates a seperate data set just for Nonsense and CIA books
CIA <- Books |>
  filter(str_detect(title, "The 2002 CIA World Factbook"))

```

*Covers the use of 1 str function, and cleaning of the data.*

**Cleaning the data using str functions**

```{r}
library(stringr)

nonsense_clean <- nonsense |>
  mutate(text = str_remove_all(text, "[[:punct:]]")) |>  #Remove all punctuation from the text column
  mutate(text = str_to_lower(text)) |> #Convert all text to lowercase to avoid treating words like "Man" and "man" as different words                   
  mutate(text = str_squish(text)) #Remove extra spaces, helps when we are tokenizing                        

CIA_clean <- CIA |>
  mutate(text = str_remove_all(text, "[[:punct:]]")) |>  
  mutate(text = str_to_lower(text)) |>  
  mutate(text = str_squish(text))

```

This covers three str functions: `str_remove_all()`, `str_to_lower()`, `str_squish()`.

## **  Tokenizing the Data:**

```{r}
tidy_nonsense <- nonsense_clean |>
  mutate(line = row_number()) |>
  unnest_tokens(word, text, token = "words")

# Tokenize CIA Factbook
tidy_CIA <- CIA_clean |>
  mutate(line = row_number()) |>
  unnest_tokens(word, text, token = "words") #tokenizes
smart_stopwords <- get_stopwords(source = "smart")

#wanted to find numbers just in case I decided to do any number analysis of the text
CIA_numbers <- tidy_CIA |>
  filter(str_detect(word, "[0-9]+")) |>
  count(word, sort = TRUE)

#also wanted to find long words in the nonsense book
nonsense_longwords <- tidy_nonsense |>
  filter(str_detect(word, "\\b[a-z]{5,}\\b")) |>
  count(word, sort = TRUE)


# Clean Nonsense words
tidy_nonsense_nostop <- tidy_nonsense |>
  anti_join(smart_stopwords, by = "word") |> #need to remove stopwords
  filter(!is.na(word)) #when words are removed there were some lines that were empty, we need to remove NAs



# Clean CIA words
tidy_CIA_nostop <- tidy_CIA |>
  anti_join(smart_stopwords, by = "word")|>
  filter(!is.na(word)) #same as above

```

To prepare my texts for the rest of the project, I used tokenizing to split the text into individual words, which makes it easier to count words and match them to sentiments. I also removed stop words and NAs to clean up the data. I used regular expressions to find numbers in the CIA Factbook and long words in The Book of Nonsense. This step helped me get the text ready for further analysis like word counts, word clouds, and sentiment graphs.

## Plotting the Top 10 words in each text:

```{r}
library(ggplot2)
library(forcats)

# Top words in Nonsense
tidy_nonsense_nostop |>
  count(word, sort = TRUE) |>
  slice_max(n, n = 10) |>
  ggplot(aes(fct_reorder(word, n), n)) +
  geom_col(fill = "skyblue") +
  coord_flip() +
  labs(title = "Top 10 Words in The Book of Nonsense (No Stopwords)",
       x = NULL, y = "Word Count")

# Top words in CIA
tidy_CIA_nostop |>
  count(word, sort = TRUE) |>
  slice_max(n, n = 10) |>
  ggplot(aes(fct_reorder(word, n), n)) +
  geom_col(fill = "red") +
  coord_flip() +
  labs(title = "Top 10 Words in The 2002 CIA World Factbook (No Stopwords)",
       x = NULL, y = "Word Count")


```

When comparing the top 10 words in The Book of Nonsense and The 2002 CIA World Factbook, clear differences emerge in both focus and language. In The Book of Nonsense, the most frequent words are centered around people, such as "man," "person," "lady," and "young," which shows the book’s focus on human characters. I found it most interesting that "man" is the top word and there is no mention of "woman"; instead, the term "lady" is used. This reflects 1800s' gender norms where "lady" was the common way to refer to women in a polite or literary context. In contrast, the CIA Factbook shows a heavy emphasis on numbers, places, and statistics, with top words like "est," "km," "2002," "population," and "rate." There are almost no references to people or emotions which shows its role as a factual and neutral government report. Overall, The Book of Nonsense is focused on characters and actions, while the CIA Factbook prioritizes data and information about countries and populations. The contrast between the two reflects both the purpose of the texts and the historical context in which they were written.

## Are there any emotions in the CIA Factbook?

Next, I will compare the emotions in The Book of Nonsense to the 2002 CIA Factbook... if there are any that is.

```{r}
nonsense_emotions <- tidy_nonsense |> #joining the nonsense data with nrc for emotions 
  inner_join(nrc, by = "word") |>
  count(sentiment, sort = TRUE)
nonsense_emotions |> #plots the emotions
  ggplot(aes(x = sentiment, y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Emotion in The Book of Nonsense", y = "Word Count", x = "Emotion") +
  theme_minimal() #plots the sentiment


CIA_emotions <- tidy_CIA |> #joining the CIA data with nrc for emotions 
  inner_join(nrc, by = "word") |>
  count(sentiment, sort = TRUE)
CIA_emotions |>
  ggplot(aes(x = sentiment, y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Emotions in The CIA Factbook", y = "Word Count", x = "Emotion") +
  theme_minimal()


```

The sentiment of the CIA Factbook is much more positive than expacted. This is surprising as it aims to be a neutral report, but I thought it was interesting that there were so many instances of emotions in the text regardless. I was expecting the chart to be near 0 on all. The Book of Nonsense has a much more uneven distribution of emotions, skewing towards negative. I found this surprising as the description of nonsense and whimsical poems sound positive. However, R's sentiment analysis found it to be the oppose. This analysis between the two texts shows how texts with much different goals are going to have a unique distribution of sentiment depending on said goals.

## Creating word clouds to compare the texts:

```{r}
library(reshape2) #needed these packages to create a word cloud
library(wordcloud)

# Wide format is needed to create word cloud
nonsense_sentiment_words <- tidy_nonsense |> 
  inner_join(get_sentiments("bing"), by = "word") |>
  count(word, sentiment, sort = TRUE)

nonsense_wide <- nonsense_sentiment_words |>
  acast(word ~ sentiment, value.var = "n", fill = 0)

# Plot comparison cloud
comparison.cloud(nonsense_wide,
                 colors = c("darkred", "gold"),
                 max.words = 100,
                 title.size = 1.5)



```

```{r}
# Wide is needed for CIA Factbook
CIA_sentiment_words <- tidy_CIA |>
  inner_join(get_sentiments("bing"), by = "word") |>
  count(word, sentiment, sort = TRUE)

CIA_wide <- CIA_sentiment_words |>
  acast(word ~ sentiment, value.var = "n", fill = 0)

# comparison cloud
comparison.cloud(CIA_wide,
                 colors = c("darkred", "gold"),
                 max.words = 100,
                 title.size = 1.5)

```

The word clouds show comparison between the positive and negative sentiments in each of the texts. The nonsense book shows that the emotions are much more personal where as the CIA book is very focused on government and societial interactions as a whole. Overall, this project was interesting because we got to see may differences between a text written for pleasure and entertainment during the mid 1800s and a text written by the government 156 years later.
